---
title: "ARIMA"
output: html_document
---

An *ARIMA* model is an *Integrated Autoregressive Moving Average Process*. It is a univariate model that is similar to ARMA but can deal with non-stationarity.
* p = AR polynomial: variable of interest is regressed on its own prior values
* d = Integrated part: Indicates that the data values have been rteplaced with the *difference* between their values and the previous values (can be more than once)
* q = MA polynomial: variable of interest depends linearly on the current and various past values of a stochastic term.

When two out of 3 terms are zeros, the model can be referred to as AR, I, or MA. E.g,
* ARIMA(1, 0, 0) = AR(1)
* ARIMA(0, 1, 0) = I(1)
* ARIMA(0, 0, 1) = MA(1)

### ARIMA(0, 1, 0)
Equivalent to a random *walk model*. The formula $X_t = X_{t-1} + \varepsilon_t$ shows that there is only one thing that helps us predict the future term, which is the past term. An ARIMA(1, 0, 1) with a constant: $X_t = c + X_{t-1} + \varepsilon_t$ is a random walk with drift.

### ARIMA(0, 0, 0)
A *White noise model. Nothing can be used to predict the future outcome.

```{r, message=FALSE}
library(FitAR)
library(forecast)
library(tseries)
library(gplots)
library(aTSA)
library(astsa)
library(lmtest)
foodcorp.df <-read.csv("SLIM0405DataSeries.csv")
```

```{r }
plot(foodcorp.df[,c(1)],foodcorp.df[,c(8)], type="l", col="red", lwd=1, xlab="weeks", ylab="sales", main="Volume sales over time")
```

### ADF test for stationarity
```{r }
adf.test(foodcorp.df[,c(8)], nlag = 4, output = TRUE)
```

### PP test for stationarity
```{r }
pp.test(foodcorp.df[,c(8)], output = TRUE)
```
### ACF
Empirical ACF decays only very slowly to zero
```{r }
foodacf = acf(foodcorp.df[,c(8)], lag.max = 10)
```

### PACF
```{r }
foodpacf = pacf(foodcorp.df[,c(8)], lag.max = 10)
```

### Differencing first-order
```{r }
foodsel = foodcorp.df[,c(8)]
fooddif = diff(foodsel)
```

### ADF test after differencing
```{r }
adf.test(fooddif, nlag = 4, output = TRUE)
```

### PP test after differencing
```{r }
pp.test(fooddif, output = TRUE)
```

### Empirical ACF after differencing
```{r }
foodacf = acf(fooddif, lag.max = 10)
```

### Empirical PACF after differencing
```{r }
foodpacf = pacf(fooddif, lag.max = 10)
```

### auto.ARIMA(1, 1, 1)
```{r }
foodsel = foodcorp.df[,c(8)]
tsfoodsel = as.ts(foodsel)
auto.arima(tsfoodsel, test = c("adf"), trace = TRUE)
```
