---
title: "Intervention_Analysis"
output: html_document
---

## __Intervention Analysis__
Suppose that at lag $t = T$ where $T$ is known, there has been some sort of intervention to a time series. An intervention can be a change to a procedure, or law, or policy. We want to estimate how much the intervention has changed the series. For example, a new speed limit is implemented and we want to know how much this new speed limit has affected accident rates.

Suppose we have the following ARIMA model with no intervention:
$$x_t = \mu + \dfrac{\Theta(B)}{\Phi(B)}\varepsilon_t$$
* $\Theta(B)$ being the MA polynomial;
* $\Phi(B)$ being the AR polynomial
* $\varepsilon$ being the error term

Now, let $z$ equal the amount of change at time $t$ that can be attributed to the intervention. By definition, $z_t = 0$ before the intervention at time $T$. However, the value $z_t$ may (or may not) be 0 after time $T$. This can be expressed in the following model:

$$x_t = \mu + z_t + \dfrac{\Theta(B)}{\Phi(B)}\varepsilon_t$$

## __Patterns for the intervention analysis__

* Permanent constant change to the mean to each value after time $T$.
* A brief constant change to the mean level after which there is no effect.
* Gradual increase or decrease to a new mean level.
* Initial change followed by gradual return to the no change state.

## __Intervention Analysis in R__ 

```{r read file, echo=FALSE, message=FALSE}
library(FitAR)
library(forecast)
library(tseries)
library(gplots)
library(aTSA)
library(astsa)
library(lmtest)
foodcorp.df <- read.csv("../SLIM0405DataSeries.csv")
```

```{r }
plot(foodcorp.df[,c(1)],foodcorp.df[,c(9)], type="l", col="red", lwd=5, xlab="weeks", ylab="sales", main="Volume sales over time", ylim = c(0,100))

```
## Step 1: Observe ACF and PACF 
However, __we will only do so for the period before the intervention__. In this case, it's up to approx. week 144.

```{r }
foodsel = foodcorp.df[,c(9)]
foodsel_x = foodcorp.df[,c(1)]
foodsel2 = foodsel[c(1:144)]
foodsel2_x = foodsel_x[c(1:144)]
foodacf = acf(foodsel2, lag.max = 10)
foodpacf = pacf(foodsel2, lag.max = 10)
```

```{r }
plot(foodsel2_x, foodsel2, type="l", col="red", lwd=5, xlab="weeks", ylab="sales", main="Volume sales over time", ylim = c(0,100))
```


From the looks of it, we seem to have an AR(1) process. There is a single spike in the PACF whilst the ACF plot is slowly decaying.

## Step 2: Observe Unit Roots
```{r }
adf.test(foodsel2)
pp.test(foodsel2)
```


## Step 3: Fit the AR(1) model using Ordinary Least Squaree regression

```{r }
tssales = as.ts(foodsel2)
# AR is based on its previous value so one lag is required
tsfoodsel3 = ts.intersect(tssales,tssales1=lag(tssales,-1))

m1 <-lm(tssales~tssales1, data=tsfoodsel3, na.action = NULL)
summary(m1)
```

## The same model can be fit using auto.arima but it gives different coefficients.
```{r }
auto.arima(tssales, test = c("adf"), trace = TRUE)
```

## Modelling the interventions

```{r }
tsfoodsel2 = as.ts(foodcorp.df)
tssales = tsfoodsel2[,c(9)]
tsintervent1 = tsfoodsel2[,c(13)]
tsintervent2 = tsfoodsel2[,c(14)]
tsfoodsel4 = ts.intersect(tssales, tssales1=lag(tssales,-1), tsintervent1, tsintervent2)

m2 <-lm(tssales~tssales1 + tsintervent1 + tsintervent2, data=tsfoodsel4, na.action = NULL)
summary(m2)
```


