---
title: "IRF"
author: "Michiel van Eldik"
date: "10/13/2020"
output: html_document
---

Once we have specified our dynamic model system. Nice! That's really the heart of what we do. We can now start with making our __estimations__. These can be used to make __policy decisions__. However, we have included so many parameters, that it may be nearly impossible to interpret all parameters.

* __Exogeneous__ is actually really straight forward
* __Endogeneous__ is not so straight-forward

If you have this whole system of feedback and lags, it's almost impossible to get what's going on.


## __IRF in practise__

The basic idea is that
    * We apply a "shock" to one of the series (e.g., advertising)
    * We trace the effect of this shock through the system on each of the other series (sales, price, competitor advertising, etc.)
The we calculate between __two series__:
    * One benchmark series where nothing happens
    * One shock series where the shock is applied.
    
    = calculate the difference between these two series.
    
What happens to our system if I apply the shock, what happens if I don't do it? What is the additional sales due to the shock in sales?


### __When variables are stationary__
If you have stationary variables...

No lasting effects of our shocks. Series return to it's pre-shock level.

Effects to analyse:
  * Immediate effect
  * Cumulative effect (in this case interesting as there is no permanent effect) i.e., what is the __full effect overall__
  * Permanent effec = 0

### __When variables are non-stationary__
If you have evolving variables...

Effects do not die over time.
Series may not return back!

    * Immediate effect
* Cumulative effect (less interesting
    
## __Restricted IRF__
Standard situation: Estimate a full VAR/VEC without parameter restrictions. __All endogeneous variables are allowed to affect each other at every single lag__ 
But how to observe the relative contribution of different variables to the final outcome observed?

Solution: Compare the restricted IRF with the full IRF to get insight on the relative contribution.


E.g., Trace the effect of advertising on sales, while keeping competitor variables constant. What is the pure effect of our advertising, without any competitor reaction effects?

## __Generalized IRF__

With the restricted IRF, you have to specify logic yourself. You impose the order yourself: Advertising, performance, satisfaction. That might have an effect. BUT __sometimes the ordering is not clear__.

There are all kinds of metrics like online purchase-funnel behaviour, online metrics, etc.

That's where we get the idea of __Generalized IRF__ = calculate the immediate effect from the variance-covariance matrix.



Standard R packages don't exist for this.



## __Case__

```{r }
library(lmtest)
library(vars)
library(aTSA)

railways.df <-read.csv("../SLIM0607Railways.csv")
```

```{r }
#Determining lag length of the endogenous variables
railendo = railways.df[,c(3,2,5)]
railexo = railways.df[,c(11,7,8,13,9,10)]    #The trend does not have to be added here, as it can be immediately added through type "both"
VARselect(railendo,lag.max = 4, type = "both", exogen = railexo)
```


```{r }
#Estimating the VAR model, and reporting results for the individual equations
railest <- VAR(railendo, p=1, type = "both", exogen = railexo)

# Pass-based variable
summary(railest,"LNKMABOFIX2")

# Ticket-based variable
summary(railest,"LNKMLOSFIX2")

# Satisfaction variable
summary(railest,"DLNAOAVERAGE_2L")
```

```{r }
#Generating the IRFs
railirf1 <- irf(railest, impulse = NULL, response = "LNKMABOFIX2", n.ahead = 12,
               ortho = TRUE, cumulative = FALSE, boot = TRUE, ci = 0.95,
               runs = 500)  
plot(railirf1)

railirf2 <- irf(railest, impulse = NULL, response = "LNKMABOFIX2", n.ahead = 12,
               ortho = TRUE, cumulative = TRUE, boot = TRUE, ci = 0.95,
               runs = 500)  
plot(railirf2)

railirf3 <- irf(railest, impulse = NULL, response = "LNKMLOSFIX2", n.ahead = 12,
               ortho = TRUE, cumulative = FALSE, boot = TRUE, ci = 0.95,
               runs = 500)  
plot(railirf3)

railirf4 <- irf(railest, impulse = NULL, response = "LNKMLOSFIX2", n.ahead = 12,
               ortho = TRUE, cumulative = TRUE, boot = TRUE, ci = 0.95,
               runs = 500)  
plot(railirf4)

railirf5 <- irf(railest, impulse = NULL, response = "DLNAOAVERAGE_2L", n.ahead = 12,
                ortho = TRUE, cumulative = TRUE, boot = TRUE, ci = 0.95,
                runs = 500)  
plot(railirf5)

railirf6 <- irf(railest, impulse = NULL, response = "DLNAOAVERAGE_2L", n.ahead = 12,
                ortho = TRUE, cumulative = TRUE, boot = TRUE, ci = 0.90,
                runs = 500) 
plot(railirf6)
```

